Optimal (Multiway) Spatial Joins

### 2024-05-26 msj.tex
yufei

This section will tackle the (generic form of the) $k$-SJ problem, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$. The join result, denoted by $\J(R_1, R_2, ..., R_k)$, is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying the condition that $B_\bm{t}$ --- defined as $\bigcap_{i=1}^k r_i \ne \emptyset$; see \eqref{eqn:B_t} --- is non-empty.

@@

Consider any result tuple $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$, and let $p$ be the top-left corner of $B_\bm{t}$. Depending on how $p$ is determined, we classify $\bm{t}$ into one of the two categories below: 
\myitems{
    \item {{\bf Cat.\ 1:}} $p$ is the top-left corner of $r_i$, for some $i \in [k]$.

    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$. Note that this means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
}
Figure~\ref{fig:ksj:cats} illustrates a tuple of each category, assuming $k=4$.

@@

We are now ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
\myeqn{
    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
}
we add $(r_1, ..., r_{k-1}, r_k)$ to $\J^\cat_k$. Performing the above for all $p \in P'$ correctly generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time. The way $P'$ is computed ensures that $\contained_{R_j}(p)$ is non-empty for all $j \in [k-1]$. Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)|$ is bounded by $\prod_{j \in [k-1]} |\contained_{R_j}(p)|$. This implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.

>>>>

This section will tackle the $k$-SJ problem in its general form, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$. The join result $\J(R_1, R_2, ..., R_k)$ is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying the condition that $B_\bm{t}$ --- which is $\bigcap_{i=1}^k r_i \ne \emptyset$ (see \eqref{eqn:B_t}) --- is non-empty.

@@

Consider any result tuple $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$, and let $p$ be the top-left corner of $B_\bm{t}$. Depending on how $p$ is determined, we classify $\bm{t}$ into one of the two categories below: 
\myitems{
    \item {{\bf Cat.\ 1:}} $p$ is the top-left corner of $r_i$ for some $i \in [k]$.

    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$. This means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
}
Figure~\ref{fig:ksj:cats} illustrates a tuple of each category, assuming $k=4$.

@@

We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
\myeqn{
    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
}
we add $(r_1, ..., r_{k-1}, r_k)$ to $\J^\cat_k$. Performing the above for all $p \in P'$ correctly generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time. The way $P'$ is computed ensures that $\contained_{R_j}(p)$ is non-empty for all $j \in [k-1]$. Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)|$ is bounded by $\prod_{j \in [k-1]} |\contained_{R_j}(p)|$. This implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.

We can now conclude that the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.
